 Casting multiple shadows: high-dimensional interactive data visualisation with tours and embeddings

### 2020-04-17 paper.Rmd
update bibliography; wrap lines
section on DR methods in progress

Embedding methods can be classified into two broad groups: linear and
non-linear methods. Linear methods perform a linear transformation of the data;
one example is principal components analysis (PCA) which performs an
eigendecomposition of the estimated sample covariance matrix. The eigenvalues
are sorted in decreasing order and represent the variance explained by each
component (eigenvector). A common approach to deciding on the number of
principal components to retain is to plot the proportion of variance explained
by each component and choose a cut-off.

Non-linear methods generally perform pre-processing on the high-dimensional data
such as generating a neighborhood graph and perform transformations on the
pre-processed form. We restrict our attention to two methods that are commonly
used in the literature: t-distributed stochastic neighbor embedding
(t-SNE) and uniform manifold alignment and projection (UMAP). The t-SNE
algorithm estimates the pairwise similarity of (Euclidean) distances of points in a high
dimensional space using a Gaussian distribution and then estimates a
configuration in the low dimensional embedding space by modelling similarities
using a t-distribution with 1 degree of freedom. The resulting configuration is
the one that minimizes the Kullback-Leibler divergence between the two
distributions. A recent theoretical contribution by @Linderman2019-dq proved
that t-SNE can recover spherical and well separated cluster shapes, and
proposed new approaches for tuning the optimisation parameters. It is a known
problem that t-SNE can have trouble recovering global structure and that
configurations can be highly dependent on how the algorithm is initialised and
parameterized [@Wattenberg2016-ji; @Kobak2019-lm]. UMAP is a method that is
related to LargeVis [@Tang2016-oz], and like t-SNE acts on the k-nearest
neighbor graph. Its main differences are that it uses a different cost function
(cross entropy) which is optimized using stochastic gradient descent and
defines a different kernel for similarities in the low dimensional space. Due
to it's computational speed it's possible to generate UMAP embeddings in more
than three dimensions.

@@

## Dimensionality Reduction

@@

#### Using t-SNE

@@

### Case Study 3: Single cell mouse retina data

>>>>

@@

## Dimensionality Reduction Overview

In the following section, we provide a brief overview of the goals
of DR methods with respect to the data analyst, in addition to their high level the 
mathematical details. To begin we suppose the data is in the form of an $n \times p$ 
rectangular matrix of real numbers, $X$, where $n$ is the number of  observations
in $p$ dimensions. The purpose of any DR algorithm is to find low-dimensonal
representation of the data, $Y$, such that $Y$ is an $n \times d$ matrix 
where $d \ll p$. The aspiration is that the DR procedure will remove noise
in the data while revealing hidden patterns and structure.

DR methods can be classified into two broad groups: linear and
non-linear methods. Linear methods perform a linear transformation of the data,
that is, $Y$ is a linear transformation of $X$;
one example is principal components analysis (PCA) which performs an
eigendecomposition of the estimated sample covariance matrix. The eigenvalues
are sorted in decreasing order and represent the variance explained by each
component (eigenvector). A common approach to deciding on the number of
principal components to retain is to plot the proportion of variance explained
by each component and choose a cut-off.

For non-linear methods $Y$ is generated via a pre-processed form of the input 
$X$ such as the $k$-nearest neighbours graph or via a kernel transformation. 
In this paper we restrict our attention to two methods that are commonly
used in the literature: t-distributed stochastic neighbor embedding (t-SNE) and
uniform manifold alignment and projection (UMAP). The t-SNE algorithm estimates
the pairwise similarity of (Euclidean) distances of points in a high
dimensional space using a Gaussian distribution and then estimates a
configuration in the low dimensional embedding space by modelling similarities
using a t-distribution with 1 degree of freedom. The resulting configuration is
the one that minimizes the Kullback-Leibler divergence between the two
distributions. A recent theoretical contribution by @Linderman2019-dq proved
that t-SNE can recover spherical and well separated cluster shapes, and
proposed new approaches for tuning the optimisation parameters. It is a known
problem that t-SNE can have trouble recovering global structure and that
configurations can be highly dependent on how the algorithm is initialised and
parameterized [@Wattenberg2016-ji; @Kobak2019-lm]. UMAP is a method that is
related to LargeVis [@Tang2016-oz], and like t-SNE acts on the k-nearest
neighbor graph. Its main differences are that it uses a different cost function
(cross entropy) which is optimized using stochastic gradient descent and
defines a different kernel for similarities in the low dimensional space. Due
to it's computational speed it's possible to generate UMAP embeddings in more
than three dimensions.

@@

@@

### Case Study 3: QuickDraw images
